A Multi-telescope Campaign on FRB 121102: Implications for the FRB Population

### 2017-07-19 referee_response.txt

General comments:

We thank the referee for a thoughtful review and suggestions for improvement. We have made almost all the recommended changes. In some cases, the intent of the comment was not clear, so we made changes while requesting clarification. In a few cases, we have responded with an explanation of why no change was made.

@@

I think this work is worthy of publication in the journal with moderate revision of the manuscript. Some moderate revision of the text is needed (see below) as well as some structural and conceptual changes. I have included my comments on the manuscript below with general large comments (1, 2, 3..) on the paper at the beginning, and major (A, B, C,...) and minor (i, ii, iii,...) comments per section.

General comments on the manuscript:

1. Throughout the text it is mentioned that this is the first simultaneous detection of an FRB pulse at multiple observatories, but I don't believe this is correct. The observations detailed in Marcote et al. 2017 include VLBI observations of individual pulses from FRB 121102 detected at different observatories. In that sense, this paper does not present the first simultaneous detection. Perhaps the authors mean first simultaneous detection at multiple radio frequencies, but if this is the case this statement should be revised throughout the manuscript.

>>> We do believe the VLA+AO burst is the first of its kind both as a multi-frequency and as a multi-telescope detection. The first VLA+Arecibo burst was detected on 17 September, while the bursts reported in Marcote et al were detected on 20 September. The publication date of the VLA+AO (discussed in Chatterjee et al 2017) and EVN bursts were simultaneous, so it is also not true that one was published first. We would prefer to keep the current wording.

@@

>>> Done.

Observations:

Figure 1: This figure is very busy and difficult to parse. I would suggest the following changes:
- Change the symbols for each observing epoch to the number of bursts detected in that observation (0, 1, 2, etc)
- Create two extra rows and plot the VLA observations at L-band, S-band, and C-band separately
- Next to the name of each telescope on the right hand side of the figure (or replacing the name of the telescope) provide the observing frequency, e.g. LWA (70 MHz), VLA (1.4 GHz), VLA (3 GHz), VLA (4.5 GHz), etc.
- Use circles to indicate overlapping frequency coverage and dispense with bold circles and dashed lines
- Use a solid line to indicate the observation where a burst was detected at both VLA and Arecibo

>>> Figure 1 is fairly dense and takes time to interpret. We must confess that it's design is limited also by our understanding of the plotting library. Some features that you suggest, such as the circles (squared ellipses?) that span multiple rows are beyond our skill to create. Would it help the current design to add a frequency label along with the telescope name on the right column?

@@

Table 1: All times should be in the format HH:MM:SS, e.g. 03:25:29
Table 1: Footnote ordering: Footnotes should be labeled in the order in which they appear, so footnote d should be relabeled as b and so on
Table 1: Footnote wording: This relates to comment #2 at the beginning of this report. I believe it would be better to label these as "detection" and "non-detection" since this is what is implied by "coverage" and the lack of a * icon indicating detection.

>>> Done. The last comment was a bit unclear to me. After rewording, we decided to use a letter footnote symbol for all nondetections and keep the * symbol to denote the Arecibo detection.

@@

Equation 1: A is typically used for a unitless amplitude, but in this case is given in units of flux density. If this value is indeed meant to be the flux density of the pulse then $S$ should be used as the symbol for flux density instead of A.

>>> Done.

A. Paragraph beginning "The parameters for the best model ... " : Your fits to the pulse spectra seem to indicate that all but two of the VLA burst detections are well-modeled by a Gaussian envelope completely contained within the 3 GHz band. How likely is this? It seems to be implying that 3 GHz was the perfect frequency at which to center your observations. While this seems fortunate it doesn't make immediate physical sense to me. I might have expected some uniformly broadband pulses, some with an envelope within your observing band, and some with an envelope that extends beyond the observing band, but you find that most are described by an entirely in-band envelope. Is this physical or could this be an artefact of the model fits you have applied to the data? If you believe that this is indeed an intrinsic effect I would like to see more discussion of the physical mechanism behind the very fortunate 3 GHz observing window.

>>> Our interpretation is that this is a detection bias and has no physical meaning. If the bursts have a peak in their spectrum, then it is easiest to detect bursts with a peak inside the observing band. Spectral peaks outside or at the edge of the band have reduced measured flux and lower significance. 
>>> It is not clear how to discuss this in the text, but we'd welcome suggestions.

@@

Table 2: Footnote labeling: The labeling here is not consistent with the labeling used in Table 1.

>>> The addition of the last footnote requires this set of footnotes to be in a different order than in Table 1. It is not clear how to include all this information with the same footnote marks as in Table 1.

@@

>>> The spectra are extracted at the best-fit (SNR maximizing) DM, which indeed varies from burst to burst. Forcing to a single DM would effectively cut some flux out, so the bursts would become narrower in frequency. In that sense, the spectrum is sensitive to the DM. Our method is intended to extract and model all flux from the burst.

3.2.2 Dispersion:

i. "There is a weak correlation between ... temporal width." : It would be useful information in this section if you could elaborate on how sensitive your observations are or aren't to substructure within each pulse that may be hindering DM fitting? For example, if there was a pulse with substructure within a single 5ms VLA time sample can you estimate the error this would propagate to the DM fit?

>>> A sentence in the previous section ("Our approach simultanesously models...") attempts to give some context for what the modeling can do. The Bayesian approach cannot say anything about the appropriateness of the model, only what can be inferred by assuming the model is correct. Substructure within the pulses is not modeled in this analysis, but one could imagine extending the modeling in that way. We will add a comment on that.
>>> On a related note, we've reworded the second to last sentence in this paragraph to distinguish between "DM" and "apparent DM".

@@

i. For clarity, please label the rates at different observing frequencies in the text as $R_{3 GHZ}$ etc.

ii. Equation 3: Similar to Equation 1, A is given in units of flux density and should therefore be represented by the symbol $S$.

iii. dN/d logE : Elsewhere in the text this is written as dN/dE. Please be consistent in notation.

>>> Done. On the last point, the distinction between dN/d logE and dN/dE was deliberate, but we tried to avoid confusing readers. The issue is that we modeled the distribution in log space, but these kinds of results are typically discussed in linear space. Since the analysis is publicly available (Jupyter notebooks on github), we wanted to maintain some reference to the details of the analysis in the text. We can't do that without redoing the analysis approach entirely, as shown in the notebook.

@@

Discussion:

4.1 Burst Spectra:

A. I do not find the explanation or investigation here into the drastic DM change to be sufficient. The DM is changing (or being measured to change) by many units and disagrees with the best-fit DM at 1.4 GHz. This disagreement between 1.4 GHz and 3 GHz is either intrinsic or extrinsic (or both, I suppose). In the intrinsic case this is of enormous interest. Although an extrinsic explanation would be less physically interesting it would nonetheless provide some helpful answers. Paragraph 6 of this section gives some attention to this issue although it doesn't address the impact that some propagation effects might have on the pulse at different radio frequencies. Given the unique multi-radio-frequency aspect of this work, such a discussion would be valuable. An additional bit of text dedicated to this topic would be illuminating.

>>> We have expanded the discussion of the burst DMs in this section. A new paragraph discusses intrinsic structure more explicitly.

@@

ii. "At the typical Parkes ... at $z \approx 1$" : Please clarify this statement. Do you mean that a Parkes survey could detect the brightest pulse from FRB 121102? The brightest X pulses? The discovery pulse (I think this one was too weak for Parkes)? The "average" pulse?

>>> Reworded. The largest burst energy detected by the VLA was 10^40, which is detectable by the typical Parkes survey at z=1.

iii. "understated by a factor of a few by ignoring bursts that fall outside the observing band" : The data presented in this paper are the best currently available to actually constrain this factor. Would it be possible to provide an estimate of this factor using the detections/non-detections from these observations?

>>> We do not believe that these data can address this question. We have identified a quasi-broadband structure for burst spectra and have shown that detection of a burst at one frequency does not predict the detection of that same burst at another frequency. As we discuss, an implication of this is that the burst rate cannot be defined from any single observing band.
>>> We are working on a comprehensive analysis of the Arecibo data (at 1.4 GHz) during this observing period. The joint consideration of the VLA and AO burst rates, along with their time and luminosity distributions, will be required to estimate this "true" burst rate.

@@

iv. "$5 \times 10^2$ and $5 \times 10^3$ times" : Is this range of repetition times sufficient for the repeater? In the observations from just the past 1 year this FRB has emitted O[100] pulses, I believe, and was active at least back to 2012, possibly much longer. Have you or others in the FRB 121102 team attempted to estimate the total number of pulses this FRB may have given off assuming this activity? In doing this does this particular source still fall in the range you've given here? If so, it is worth mentioning where you believe FRB 121102 falls in this range.

>>> Good point. We've split this paragraph in two and added some discussion on this point. We've also added a sentence to the conclusion on this point.

### 2017-07-19 repeater-vladata.tex
through comments. sending to team

\subsection{Multi-Observatory Burst Spectrum}

@@

\subsection{VLA Bursts}

@@

\section{discussion}
\label{sec:disc}
\subsection{Burst Spectra}

@@

If \frb\ burst spectra are typical of the larger FRB population, then population models need to be generalized beyond the assumption of a spectral power law. Most obviously, \frb\ implies that future multi-telescope searches for bursts are unlikely to simultaneously detect bursts in different bands \citep[c.f.][]{1999ApJ...517..460S}. Second, since burst spectra have limited bands, the burst detection rate at one frequency may not represent that at another. Rate estimates for the FRB population will need to explicitly account for the frequency-dependent rate. Finally, we note that the odds of detecting a burst will improve with bandwidth beyond the gain in sensitivity, since bandwidths wider than the FRB characteristic width are more likley to cover the burst envelope. However, FRB search algorithms will need to be modified to search for bursts with spectral width less than the full bandwidth.

After correcting for barycentric and dispersion delays, the VLA+AO burst spectrum has a residual temporal drift. This drift can be interpreted as an excess DM, but contemporaneous observations at 1.4~GHz measured a significantly lower DM (560.5 pc cm$^{-3}$; Hessels et al, in prep). This excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016Natur.531..202S,2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen simultaneously from 1.4 to 3~GHz.

One interpretation for the burst-dependent DM is that the bursts have a frequency-dependent pulse shape. Pulsar pulse shapes evolve with frequency on GHz frequency scales, presumably due to changes in the beam shape \citep{1988MNRAS.234..477L}. The burst-dependent DM changes are on the order of 1\% of the total DM (equivalent to a delay rate up to 2~ms GHz$^{-1}$), which is much larger than typically observed \citep{2017MNRAS.466.3706L}. Our modeling of the VLA burst dynamic spectra show a weak correlation between larger apparent DM and pulse width. This could be a signature of unresolved spectrotemporal structure in the VLA burst spectra that biases the measured DM.

@@

\citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered by its magnetic field as $\approx3000 (0.1/\Omega_b)$. This rate comparison suggests that the SLSN-I rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate. Under that model, \frb\ could emit bursts for roughly 400 days at the peak burst rate measured by the VLA. Continued monitoring of \frb\ bursts will continue to be useful in testing models for FRB origin.

@@

Assuming that \frb\ is representative of the FRB population, we calculate a volumetric birth rate of FRB sources that does not depend on telescope sensitivity. We estimate a volumetric rate of FRB sources of $R_{\rm{FRB}} \approx 5\times10^{-5} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$, which is appropriate for \frb-like burst energies of $10^{40}$\ erg that are detectable out to $z=1$. This rate is broadly consistent with models of FRBs from young pulsars or magnetars born in SLSN-I or LGRB, if the typical FRB repeats on the order of $10^3$\ times over its lifetime.

@@

%\bibliographystyle{yahapj}
\bibliographystyle{apj}

>>>>

Here we present the analysis of the VLA bursts detected toward \frb\ with a focus on constraints from coordinated radio observations. In total, nine bursts were detected by the VLA. Three of those bursts had observing coverage by other telescopes, yet were not detected. One of the VLA bursts was detected by another telescope, Arecibo. In \S \ref{sec:vlaao}, we describe the spectral properties implied by multi-frequency detections and nondetections of VLA bursts at 3~GHz. Section \ref{sec:vla} presents a deeper study of the VLA bursts alone, including a spectrotemporal modeling of the bursts, their energies, and their temporal distribution.

\subsection{Multi-Observatory Burst Spectrum}
\label{sec:vlaao}

@@

\subsection{VLA Bursts}
\label{sec:vla}

@@

\section{discussion}
\label{sec:disc}

\subsection{Burst Spectra}

@@

If \frb\ burst spectra are typical of the larger FRB population, then population models need to be generalized beyond the assumption of a spectral power law. Most obviously, \frb\ implies that future multi-telescope searches for bursts are unlikely to simultaneously detect bursts in different bands \citep[c.f.][]{1999ApJ...517..460S}. Second, since burst spectra have limited envelopes, the burst detection rate at one frequency may not represent that at another. Rate estimates for the FRB population will need to explicitly account for the frequency-dependent rate. Finally, we note that the odds of detecting a burst will improve with bandwidth beyond the gain in sensitivity, since bandwidths wider than the FRB characteristic width are more likley to cover the burst envelope. However, to benefit from this effect, FRB search algorithms will need to search for bursts with spectral width less than the full bandwidth.

After correcting for barycentric and dispersion delays, the VLA+AO burst spectrum has a residual temporal drift. This drift can be interpreted as dispersion in excess of the nominal value inferred from contemporaneous observations at 1.4~GHz (560.5 pc cm$^{-3}$; Hessels et al, in prep). However, the VLA+AO burst is not the only one with an apparent excess DM; most burst DM measurements are inconsistent with the nominal DM (Figure \ref{fig:dmdt}). The scale of this excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016Natur.531..202S,2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen simultaneously from 1.4 to 3~GHz.

One interpretation for the burst-dependent DM is that the bursts have a frequency-dependent pulse shape. Pulsar pulse shapes evolve with frequency on GHz frequency scales, presumably due to changes in the pulsar beam shape \citep{1988MNRAS.234..477L}. The burst-dependent DM changes are on the order of 1\% of the total DM (equivalent to a delay rate up to 2~ms GHz$^{-1}$), which is much larger than typically observed for pulsars \citep{2017MNRAS.466.3706L}.

Another possibility is that there is intrinsic structure to the pulses from \frb. Figure \ref{fig:corner} shows how the spectrotemporal distribution of one burst was not well modeled by a single component. This is consistent with more complex burst structure seen toward \frb\ and other FRBs \citep{2016Natur.531..202S,2016MNRAS.460L..30C}. Our modeling of the VLA burst dynamic spectra show a weak correlation between larger apparent DM and pulse width. This could be a signature of unresolved spectrotemporal structure in the VLA burst spectra that biases the measured DM. If so, then the systematic measurement of higher DMs by the VLA ($\sim5$~pc cm$^{-3}$ higher than nominal) requires that the intrinsic structure evolve toward lower frequencies during the burst.

@@

\citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered by its magnetic field as $\approx3000 (0.1/\Omega_b)$ for an isotropic energy of $10^{40}$~erg. This rate comparison suggests that the SLSN-I rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate. The VLA observations of \frb\ imply that roughly 200 bursts of this energy were emitted during the momth of peak activity. This suggests that \frb\ data could be in tension with the young magnetar model, if this level of activity were sustained for roughly a year.

@@

Assuming that \frb\ is representative of the FRB population, we calculate a volumetric birth rate of FRB sources that does not depend on telescope sensitivity. We estimate a volumetric rate of FRB sources of $R_{\rm{FRB}} \approx 5\times10^{-5} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$, which is appropriate for \frb-like isotropic burst energies of $10^{40}$\ erg that are detectable out to $z=1$. This rate is broadly consistent with models of FRBs from young pulsars or magnetars born in SLSN-I or LGRB, if the typical FRB repeats on the order of $10^3$\ times over its lifetime. However, if \frb\ continues to produce energetic bursts at a high rate, it may argue that extrinsic effects amplify the bursts \citep{2017arXiv170306580C}.

@@

\bibliographystyle{apj}
